Last glacial cycle glacier erosion potential in the Alps

### 2021-01-24 articles/erosion/alps-erosion.tex
Draft results interpretation. On the move again.

    It is often believed that glaciers are more erosive during deglaciations
    because of higher water availability. However, surface meltwater is absent
    from the model. Higher erosion rates can be explained by enhanced ice
    dynamics after longer-term ice cover.

    The results also indicate that if a square-rule erosion law is valid, much
    of the intra-Alpine erosional landscape do not date from stages of major
    glaciation.

>>>>

    Paradoxically, the model results do not show an increase of modelled
    glacier erosion rates with increasing ice volume.  On the opposite, periods
    of ice advance and maximum glaciation correspond to comparatively slow
    erosion. While this result may seem paradoxical, it is in fact
    compatible with previous field-based studies that have evidence an inverse
    relationship between air temperatures and glacier erosion rates
    \citep{Koppes.etal.2015, Cook.etal.2020}.

    Deglaciation periods, on the other hand, which are characterised by warmer
    climate and above-balance glacier extent, yield peak domain-integrated
    erosion rates. Field-based studies have associated deglaciation periods to
    higher rates of glacial erosion and proglacial sedimentation, but the increase
    has often been attributed to higher meltwater availability, promoting river
    erosion, enhanced glacier sliding, and hydrofracturing (plucking) \aref. It
    is important to note, however, that such processed are not accounted for in
    the present study. In particular, deglaciation surface meltwater is assumed
    to instantly exit the glacier with no effect on glacier dynamics.

    The enhanced glacier dynamics during periods of ice retreat is here
    modelled to result from changes of air temperature alone, reported through
    diffusion and advection through the ice, thereby facilitating shear strain
    and eventually activating basal sliding. The observed feedbacks between
    surface meltwater and basal sliding would presumably enhance erosion peaks
    modelled to occur during deglaciation. Enhanced glacier erosion can be
    expected to accompany the ongoing 21st-century deglaciation phase.

    The results also indicate that during stages of extensive glaciation, much
    of the intra-Alpine region experiences insignificant erosion rates. For
    instance, from 35 to 18\,ka BP, terrain with elevations above 1000\,m are
    modelled to experiences virtually no glacier erosion. This would imply that
    the spectacular glacial valleys found throughout the northwestern Alps do
    not date from the LGM or similar periods, but from periods of intermediate,
    and likely shrinking, ice cover.
